Mapping stellar surfaces\\ III: A Bayesian Approach to Doppler Imaging

### 2021-10-12 src/ms.tex
final tweaks

\begin{abstract}
    We derive a novel, efficient, and closed form solution to the problem of Doppler imaging of stellar surfaces from high resolution spectral timeseries datasets.
    %
    Our model for the observed spectrum is linear in the parameters describing the specific intensity distribution on the surface, which makes the posterior distribution over surface maps computationally trivial to evaluate under certain choices of priors.
    %
    Our model is also linear in the local (rest frame) stellar spectrum, which may itself be spatially variable; this allows one to perform Doppler imaging with limited knowledge of the local stellar spectrum and therefore works on blended lines, regions of the spectrum where line formation mechanisms are not well understood, or stars whose spots have intrinsically different spectra from the rest of the photosphere.
    %
    We implement the model within the open source \starry stellar modeling framework, making it fast, differentiable, and easy to use in both optimization and posterior inference settings.
    %
    As a proof-of-concept, we use our model to infer the surface map of the brown dwarf Luhman 16B, finding close agreement with the solution of \citet{Crossfield2014}.
    %
    This paper was prepared using the \showyourwork scientific article workflow, which exposes the data, source code, and package dependencies for all of the paper's figures, allowing readers to reproduce the results in this paper exactly and with minimal effort.
\end{abstract}

@@

First, Doppler imaging is usually an ill-posed problem, meaning there is no unique solution for the surface map given a spectroscopic dataset.
This issue is typically well understood \citep[e.g.,][]{Piskunov1990,Rice2002}, and most studies circumvent this issue by employing restrictive priors to ensure the uniqueness of the solution. 
These include maximum entropy regularization \citep{Vogt1987} and Tikhonov regularization \citep[e.g.,][]{Piskunov1990}, which typically enforce sparsity and smoothness in the solution, respectively.
However, while useful for obtaining a maximum likelihood solution, both choices can lead to significant bias in the inferred map, as they can both suppress features and introduce spurious artifacts.
In fact, a recent comparison of Doppler imaging, photometric light curve inversion, and interferometry revealed qualitatively diffferent maps of the the same star \citep{Roettenbacher2017}, due to the various degeneracies afflicting each method.
%
Second, the Doppler imaging technique is generally more sensitive to longitudinal than latitudinal variations, often leading to spurious, latititudinally elongated features in the maximum likelihood solution \citep[e.g.,][]{Unruh1995,Crossfield2014b}.
%
Third, the Doppler imaging technique requires the availability of accurate spectral templates for the local rest frame spectrum of the star.
Several studies have shown that errors in these templates or the presence of unresolved line blends in the observed spectrum can introduce signficant artifacts in the map solution \citep[e.g.,][]{Rice1989,Unruh1995}.
Moreover, changes in the local spectrum over the surface of the star (such as different line depths or the presence of different absorption lines within spots) can similarly bias the inferred map if these are not properly modeled.
While some past studies have solved for variations in the local line profiles \citep{Khokhlova1976,Goncharskii1977,Goncharskii1982}, these usually make the assumption that the intensity variation is small across the surface of the star; that is, they do not simultaneously solve for the Doppler map of the surface brightness variations.
%
Fourth, and finally, Doppler imaging results reported in the literature use a variety of custom data analysis codes, most of which are not open source 
\citep[][to name a few]{Collier1997,Hackman2001,Carroll2012,Rosen2015,AsensioRamos2021}; % DOTS, INVERS7-PD, iMAP, INVERSLSD, ??
many studies also utilize data that is often not included in the resulting publication or made easily available online. 
This makes it difficult to reproduce results and compare Doppler images of different stars on the same footing. 
As argued above, detailed choices in a Doppler imaging analysis will have a significant effect on the reported point-estimate result, so this inability to compare uniformly analyzed observations is a significant hindrance to interpretation.

In this paper, the third in a series dedicated to mapping stellar surfaces \citep{Luger2021a,Luger2021b}, we seek to address these issues by developing a novel algorithm for Doppler imaging.
The various pitfalls of Doppler imaging outlined above necessitate an algorithm that is efficient enough to be used in a Bayesian posterior inference context, where one can formally address the degeneracies at play and quantify their effect on the uncertainties of the inferred stellar maps.
They also necessitate an algorithm that is flexible, capable of learning deviations from the template or capturing spatial variations in the rest-frame spectrum.
And most of all, they necessitate a user-friendly, well-documented, and open source implementation, which is sorely lacking the literature.
Our algorithm addresses these various issues by building on the work of \citet{Luger2019}, who derived analytic solutions to the light curve mapping problem by expanding the stellar surface map in the spherical harmonic basis.
While spherical harmonics are routinely used in Doppler imaging problems \citep{Deutsch1958,Deutsch1970,Falk1974,Donati2006,Kochukhov2014}, we present a new closed-form solution to the forward model for Doppler imaging when working in this basis, making our model both precise and computationally efficient.
We also expose a fundamental linearity of the model in terms of both the surface map and the rest frame spectrum, allowing us to learn both using a fast bilinear solver.
Finally, we implement our algorithm within the open source \starry code package with accompanying documentation, examples, and tutorials.\footnote{\url{https://github.com/rodluger/starry}}
All of the figures in this paper are also open source; clickable \GitHubIconBlue icons next to each figure link to the exact version of the \texttt{Python} script used to generate them.
To ensure the full reproducibility of this paper, we prepared it using the \showyourwork workflow\footnote{\url{https://github.com/rodluger/showyourwork}}, which exposes the datasets, source code, and all package dependencies needed to exactly reproduce the results.

The paper is organized as follows: in \S\ref{sec:the_problem}~and~\S\ref{sec:the_solution} we introduce the math behind the Doppler imaging problem and derive a closed form expression for the model. 
In \S\ref{sec:linear},~\S\ref{sec:inverse},~and~\S\ref{sec:bellswhistles} we demonstrate how to re-express the model as a linear operation on the input spectrum and stellar map and derive closed form expressions for the two conditioned on the data and priors.
In \S\ref{sec:spotstar} we apply our techniques to a mock problem, and in \S\ref{sec:luhman16b} we apply it to a real spectral dataset of the brown dwarf Luhman 16B.
We further discuss our results in \S\ref{sec:discussion}~and~our implementation within the \starry framework in \S\ref{sec:starry}.
We present a summary of the paper in \S\ref{sec:conclusions}.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=0.65\linewidth]{figures/compare.pdf}
        \caption{%
            The observed spectrum when the spot is in view computed with our algorithm (blue) and numerically (orange). 
            The problem setup is the same as that in Figure~\ref{fig:spot}. 
            The bottom panel shows the absolute value of the difference between the two spectra for different number of grid points in the numerical solution.
            As the number of grid points increases, the numerical solution approaches our solution.
        }
        \label{fig:compare}
    \end{centering}
\end{figure}

@@

Given this new equation, we must change how we approach the inverse problem (\S\ref{sec:inverse}) if our data vector is the normalized spectral timeseries $\fnorm$. 
When solving for the spectrum given the spherical harmonic vector (\S\ref{sec:solve_s}), we need only make the replacements 
%
$\bvec{f} \rightarrow \fnorm \circ (\bvec{B}\, \bvec{a})$ 
%
and 
%
$\boldsymbol{\Sigma}_\mathbf{f} \rightarrow \boldsymbol{\Sigma}_{\fnorm} \circ (\bvec{B}\, \bvec{a})^2$ 
%
in Equations~(\ref{eq:inverse:shat}) and (\ref{eq:inverse:scov}), where $\circ$ denotes the elementwise product.
However, when solving for the spherical harmonic vector given the spectrum (\ref{sec:solve_y}), the problem is no longer linear in $\bvec{y}$, as the vector of interest appears in both the numerator and the denominator:
%
\begin{align}
    \label{eq:norm:fy}
    \fnorm
     & =
    \frac{
        \bvec{D}
        \,
        \bvec{S}
        \,
        \bvec{y}
    }{
        \bvec{B}
        \,
        \bvec{S}
        \,
        \bvec{y}
    }
    \quad.
\end{align}
%
Unfortunately, this equation does not admit a closed form solution for $\mathbf{y}$. 
One can, however, still solve for the maximum a posteriori value of $\mathbf{y}$ in an efficient manner by iteratively solving for the baseline and then for the spherical harmonic coefficients.
We describe a procedure for this in \S\ref{sec:spotstar} below.

@@

In the following sections we discuss various approaches to solving the Doppler imaging problem for this dataset.
Unless otherwise noted, in all experiments we adopt a Gaussian prior for the spherical harmonic coefficients $\bvec{y}$ with mean $\boldsymbol{\mu}_\bvec{y} = \left(1 \quad\quad \bvec{0}\right)$ and covariance $\boldsymbol{\Lambda}_\bvec{y} = 10^{-4} \bvec{I}$, corresponding to a flat power spectrum prior.
When solving for the rest frame spectrum, we also place a Gaussian prior on $\bvec{s}$ with unit mean and covariance $\boldsymbol{\Lambda}_\bvec{s} = 10^{-3}\bvec{I}$.
These choices are arbitrary and primarily based on trial-and-error given knowledge of the true solutions.
We discuss strategies for choosing optimal priors for real datasets in \S\ref{sec:discussion:priors} below.

@@

In our inference step, we again assume we have no knowledge of the map, the baseline, or either of the spectral components.
While our bilinear method (\S\ref{sec:spot_y1bs}) could in principle work here, we find that in practice it is very difficult to converge on the global optimum without either good priors or a good initial guess.
We therefore optimize our model using the nonlinear gradient-descent \texttt{Adam} optimizer \citep{Adam} within the \texttt{pymc3} probabilistic modeling framework \citep{Salvatier2016}.
Since we are no longer restricted to using Gaussian priors on the variables (a requirement of our bilinear method), we place a unit-mean Laplacian (sparsity) prior on the two spectra with scale parameter $b = 10^{-3}$ and bounded to the range $[0, 1]$.
%
We further place a Gaussian process (GP) prior on the spectra to enforce smoothness, adopting a squared exponential kernel with amplitude $10^{-1}$ and lengthscale $10^{-2}$~nm.
%
We also solve for a single scale parameter $r \sim \mathrm{Uniform}(0, 0.75)$, equal to the ratio of the continuum levels of the two spectra (whose true value is 0.5).
Finally, in order to enforce that the two component surface maps sum to unity (as described above), we optimize for the actual \emph{pixel} intensities of the photosphere map, on which we place a simple uniform prior on $[0, 1]$; the spot map is then simply the complement.
The spherical harmonic coefficients are then computed via a fast, linear spherical harmonic transform (SHT) operation at each step, as in \citep{Bartolic2021}.
In order to satisfy the Nyquist criterion, our pixel grid consists of twice as many pixels as spherical harmonic coefficients.

@@

\section{Application to Luhman 16B}
\label{sec:luhman16b}
%

Having shown that our model performs well on synthetic data, we now turn to a real dataset: the VLT/CRIRES observations of Luhman 16B (WISE 1049-5319B) presented in the Doppler imaging study of \citet{Crossfield2014}.
Luhman 16B belongs to a binary brown dwarf system, which, at a distance of 2 pc, consists of the two closest known brown dwarfs to the Solar system.
Despite its proximity, Luhman 16B is intrinsically dim and, with a projected rotational velocity $v \sin i = 26.1$ km/s, is at the slow end of what can be imaged using the traditional Doppler imaging technique.
Therefore, unlike those of classical Doppler imaging targets, the spectral lines of Luhman 16B do not individually show large scale spot modulation (Figure~\ref{fig:spot}) from one epoch to the next; nor are the spectral templates for brown dwarfs particularly well understood.
Nonetheless, we chose to model Luhman 16B specifically to show the power behind our new methodology, in particular its ability to learn both the surface map and the rest frame spectrum simultaneously from the entire dataset---not just from unblended spectral lines that have reliable templates.

In the original Doppler imaging study of this target, \citet{Crossfield2014} took 56 spectra of each object (the angular resolution was high enough to obtain individual spectra of both brown dwarfs) over the four CRIRES detectors in the range $2.288-2.345 \mu$m over a time span of 5 hours, with individual exposures of 300 s.
The 5 hour baseline was enough to capture one full rotation of Luhman 16B, whose period was constrained to be about 4.9 hours from its photometric variability \citep{Gillon2013}.
These spectra were then downbinned to 14 epochs to boost the SNR, corrected for tellurics and spectrograph systematics, and continuum-normalized.
Each of the 14 spectra consist of 1,024 intensity measurements in each of 4 detectors, for a total of $57,344$ data points.
For more details on the data reduction, refer to \citet{Crossfield2014}.

@@

In order to produce Doppler images of Luhman 16B, \citet{Crossfield2014} used brown dwarf template spectra from the BT-Settl library \citep{Allard2013}, finding that the spectrum with $\log_{10} g = 5.0$ and $T_\mathrm{eff} = 1450$ K provided the best match to the dataset.
This spectrum was then Doppler-shifted by the systemic radial velocity of the brown dwarf and broadened to match the spectrograph resolution.
This final template spectrum is shown as the blue curves in Figure~\ref{fig:luhman16b:template_spectra}, where we again masked regions where data was discarded.

@@

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/luhman16b_data_model.pdf} % 
        \caption{%
            Calibrated and tellurics-corrected CRIRES spectra of Luhman 16B from \citet{Crossfield2014} taken at 14 epochs spanning a full rotation of the brown dwarf (black points).
            Our maximum \emph{a posteriori} solution is shown as the orange curves.
        }
        \label{fig:luhman16b:data_model}
    \end{centering}
\end{figure}
%
\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/luhman16b_spectra.pdf} % 
        \caption{%
            Unbroadened template spectra used to model the data in \citet{Crossfield2014} (blue curves). 
            These were used as a prior on the spectra inferred in this work (orange curves). 
            We infer a slightly different depth for some lines, as well as subtle changes in the continuum.
        }
        \label{fig:luhman16b:template_spectra}
    \end{centering}
\end{figure}

@@

As in \S\ref{sec:twospec}, our choice of parametrization and priors precludes the use of the fast bilinear approach outlined in \S\ref{sec:inverse}.
We therefore follow the same approach as in \S\ref{sec:twospec}, optimizing our log-likelihood function using \texttt{Adam} with a learning rate of $10^{-3}$.
Since we have a very good guess for the template spectrum, this optimization is much faster than in \S\ref{sec:twospec}, converging within $1,000$ steps.
Because of our efficient gradient-based implementation (\S\ref{sec:starry}), the entire optimization takes only about 3 minutes of runtime on a typical Macbook laptop.

\begin{figure}[t!]
    \begin{centering}
        \begin{minipage}[t]{0.49\linewidth}
            \centering
            \quad\quad\quad
            \textbf{Luger et al. (2021)}
        \end{minipage}
        \vspace{1em}
        \begin{minipage}[t]{0.49\linewidth}
            \centering
            \textbf{Crossfield et al. (2014)}
        \end{minipage}
        \includegraphics[width=0.495\linewidth]{figures/luhman16b_map.pdf} % 
        \includegraphics[width=0.462\linewidth]{static/luhman16b_crossfield.png} % 
        \caption{%
            Maximum \emph{a posteriori} surface maps of Luhman 16B inferred from the spectra in Figure~\ref{fig:luhman16b:data_model}. 
            \emph{Left:} surface map inferred using the algorithm developed in this work. 
            \emph{Right}: surface map inferred in \citet{Crossfield2014}, reproduced from Figure~2 in that paper.
            We infer the presence of a bright circumpolar spot, as well as the presence of a few dark features, most of which agree with the map presented in the original work.
        }
        \label{fig:luhman16b:maps}
    \end{centering}
\end{figure}

@@

The right panel of Figure~\ref{fig:luhman16b:maps} shows our principal result: the inferred solution for the surface brightness map of Luhman 16B; compare to the right panel, showing the solution inferred by \citet{Crossfield2014}.
Broadly speaking, the two surface maps agree rather well: we infer the same circumpolar bright spot, of about the same size and at roughly the same phase, although the exact latitude is different.
We also infer a similar dark feature on the ``back'' side of the object (visible near the sub-observer point at $t = 2.4$ hours).
Bright equatorial regions also broadly match between both solutions.
There are some important differences, such as the behavior of the $t = 2.4$ hour dark region beyond $60^\circ$ of the visible pole.
In general, we find that different choices for our priors---in particular the regularization on the pixels and the template spectrum, the prior on the limb darkening coefficient, and the value of the inclination of the star---result in maps that differ somewhat.
We do, however, agree with \citet{Crossfield2014} in that the polar bright spot is the most robust of the inferred features, followed by the bottom portion of the dark spot seen at $t = 2.4$ hours.
Once again, since this is simply an \emph{optimization}, we are unable to place reliable uncertainties on our solutions.
These require a sampling scheme like MCMC (\S\ref{sec:discussion:priors}).

@@

\subsection{Luhman 16B}
\label{sec:discussion:luhman}

>>>>

\begin{abstract}
    %
    The study of stellar surfaces can reveal information about the chemical composition, interior structure, and magnetic properties of stars.
    It is also critical to the detection and characterization of extrasolar planets, in particular those targeted in extreme precision radial velocity (EPRV) searches, which must contend with stellar variability that is often orders of magnitude stronger than the planetary signal.
    %
    One of the most successful methods to map the surfaces of stars is Doppler imaging, in which the presence of inhomogeneities is inferred from subtle line shape changes in high resolution stellar spectra.
    %
    \textbf{In this paper, we present a novel, efficient, and closed-form solution to the problem of Doppler imaging of stellar surfaces.}
    %
    Our model explicitly allows for incomplete knowledge of the local (rest frame) stellar spectrum, allowing one to learn differences from spectral templates while simultaneously mapping the stellar surface. 
    It therefore works on blended lines, regions of the spectrum where line formation mechanisms are not well understood, or stars whose spots have intrinsically different spectra from the rest of the photosphere.
    %
    We implement the model within the open source \starry stellar modeling framework, making it fast, differentiable, and easy to use in both optimization and posterior inference settings.
    %
    As a proof-of-concept, we use our model to infer the surface map of the brown dwarf WISE 1049-5319B, finding close agreement with the solution of \citet{Crossfield2014}.
    %
    We also discuss Doppler imaging in the context of EPRV studies and describe an interpretable spectral-temporal Gaussian process for stellar spectral variability that we expect will be important for EPRV exoplanet searches.
    %
    This paper was prepared using the \showyourwork scientific article workflow, which exposes the data, source code, and package dependencies for all of the paper's figures, allowing readers to reproduce the results in this paper exactly and with minimal effort.
\end{abstract}

@@

\textbf{First}, Doppler imaging is usually an ill-posed problem, meaning there is no unique solution for the surface map given a spectroscopic dataset.
This issue is typically well understood \citep[e.g.,][]{Piskunov1990,Rice2002}, and most studies circumvent this issue by employing restrictive priors to ensure the uniqueness of the solution. 
These include maximum entropy regularization \citep{Vogt1987} and Tikhonov regularization \citep[e.g.,][]{Piskunov1990}, which typically enforce sparsity and smoothness in the solution, respectively.
However, while useful for obtaining a maximum likelihood solution, both choices can lead to significant bias in the inferred map, as they can both suppress features and introduce spurious artifacts.
In fact, a recent comparison of Doppler imaging, photometric light curve inversion, and interferometry revealed qualitatively diffferent maps of the the same star \citep{Roettenbacher2017}, due to the various degeneracies afflicting each method.
%
\textbf{Second}, the Doppler imaging technique is generally more sensitive to longitudinal than latitudinal variations, often leading to spurious, latititudinally elongated features in the maximum likelihood solution \citep[e.g.,][]{Unruh1995,Crossfield2014b}.
%
\textbf{Third}, the Doppler imaging technique requires the availability of accurate spectral templates for the local rest frame spectrum of the star.
Several studies have shown that errors in line templates or the presence of unresolved line blends in the observed spectrum can introduce signficant artifacts in the map solution \citep[e.g.,][]{Rice1989,Unruh1995}.
Moreover, changes in the local spectrum over the surface of the star (such as different line depths or the presence of different absorption lines within spots) can similarly bias the inferred map if these are not properly modeled.
While some past studies have solved for variations in the local line profiles \citep{Khokhlova1976,Goncharskii1977,Goncharskii1982}, these usually make the assumption that the intensity variation is small across the surface of the star; that is, they do not simultaneously solve for the Doppler map of the surface brightness variations.
%
\textbf{Fourth}, and finally, Doppler imaging results reported in the literature use a variety of custom data analysis codes, most of which are not open source 
\citep[][to name a few]{Collier1997,Hackman2001,Carroll2012,Rosen2015,AsensioRamos2021}; % DOTS, INVERS7-PD, iMAP, INVERSLSD, ??
many studies also utilize data that is often not included in the resulting publication or made easily available online. 
This makes it difficult to reproduce results and compare Doppler images of different stars on the same footing. 
As argued above, detailed choices in a Doppler imaging analysis will have a significant effect on the reported point-estimate result, so this inability to compare uniformly analyzed observations is a significant hindrance to interpretation.

In this paper, the third in a series dedicated to mapping stellar surfaces \citep{Luger2021a,Luger2021b}, we develop a novel algorithm for Doppler imaging that addresses each of these issues.
The various pitfalls of Doppler imaging outlined above necessitate an algorithm that is efficient enough to be used in a Bayesian posterior inference context, where one can formally address the degeneracies at play and quantify their effect on the uncertainties of the inferred stellar maps.
They also necessitate an algorithm that is flexible, capable of learning deviations from the template or capturing spatial variations in the rest-frame spectrum.
And most of all, they necessitate a user-friendly, well-documented, and open source implementation, which is currently lacking the literature.
Our algorithm addresses these various issues by building on our previous work \citep{Luger2019}, where we derived analytic solutions to the light curve mapping problem by expanding the stellar surface map in the spherical harmonic basis.
While spherical harmonics are routinely used in Doppler imaging problems \citep{Deutsch1958,Deutsch1970,Falk1974,Donati2006,Kochukhov2014}, we present a new closed-form solution to the forward model for Doppler imaging when working in this basis, making our model both precise and computationally efficient.
We also expose a fundamental linearity of the model in terms of both the surface map and the rest frame spectrum, allowing us to learn both using a fast bilinear solver.
Finally, we implement our algorithm within the open source \starry code package with accompanying documentation, examples, and tutorials.\footnote{\url{https://github.com/rodluger/starry}}
All of the figures in this paper are also open source; clickable \GitHubIconBlue icons next to each figure link to the exact version of the \texttt{Python} script used to generate them.
To ensure the full reproducibility of this paper, we prepared it using the \showyourwork workflow\footnote{\url{https://github.com/rodluger/showyourwork}}, which exposes the datasets, source code, and all package dependencies needed to exactly reproduce the results.

The paper is organized as follows: in \S\ref{sec:the_problem}~and~\S\ref{sec:the_solution} we introduce the math behind the Doppler imaging problem and derive a closed-form expression for the model. 
In \S\ref{sec:linear},~\S\ref{sec:inverse},~and~\S\ref{sec:bellswhistles} we demonstrate how to re-express the model as a linear operation on the input spectrum and stellar map and derive closed-form expressions for the two conditioned on the data and priors.
In \S\ref{sec:spotstar} we apply our techniques to a mock problem, and in \S\ref{sec:luhman16b} we apply it to a real spectral dataset of the brown dwarf WISE 1049-5319B.
We further discuss our results in \S\ref{sec:discussion}, and in \S\ref{sec:gp} we focus on an immediate extension of our work: the development of an interpretable spectral-temporal Gaussian process for stellar spectral variability.
Finally, we discuss the implementation of our Doppler imaging algorithm within the \starry framework in \S\ref{sec:starry}, and present a summary of the paper in \S\ref{sec:conclusions}.

@@

\begin{figure}[t!]
    \begin{centering}
        \includegraphics[width=0.65\linewidth]{figures/compare.pdf}
        \caption{%
            The observed spectrum when the spot is in view computed with our algorithm (blue) and numerically (orange). 
            The problem setup is the same as that in Figure~\ref{fig:spot}. 
            The bottom panel shows the absolute value of the difference between the two spectra for different number of grid points in the numerical solution.
            As the number of grid points increases, the numerical solution approaches our solution.
            Note that here we assume the stellar surface can be expressed exactly as an $l_\mathrm{max} = 20$ spherical harmonic decomposition.
            When our algorithm is applied to real data, there will be errors due to the truncation of the expansion at finite spherical harmonic degree.
        }
        \label{fig:compare}
    \end{centering}
\end{figure}

@@

Given this new equation, we must change how we approach the inverse problem (\S\ref{sec:inverse}) if our data vector is the normalized spectral timeseries $\fnorm$. 
When solving for the spectrum given the spherical harmonic vector (\S\ref{sec:solve_s}), we need only make the replacements 
%
$\bvec{f} \rightarrow \fnorm \circ (\bvec{B}\, \bvec{a})$ 
%
and 
%
$\boldsymbol{\Sigma}_\mathbf{f} \rightarrow \boldsymbol{\Sigma}_{\fnorm} \circ (\bvec{B}\, \bvec{a})^2$ 
%
in Equations~(\ref{eq:inverse:shat}) and (\ref{eq:inverse:scov}), where $\circ$ denotes the elementwise product.
However, when solving for the spherical harmonic vector given the spectrum (\ref{sec:solve_y}), the problem is no longer linear in $\bvec{y}$, as the vector of interest appears in both the numerator and the denominator:
%
\begin{align}
    \label{eq:norm:fy}
    \fnorm
     & =
    \frac{
        \bvec{D}
        \,
        \bvec{S}
        \,
        \bvec{y}
    }{
        \bvec{B}
        \,
        \bvec{S}
        \,
        \bvec{y}
    }
    \quad.
\end{align}
%
Unfortunately, this equation does not admit a closed-form solution for $\mathbf{y}$. 
One can, however, still solve for the maximum \emph{a posteriori} value of $\mathbf{y}$ in an efficient manner by iteratively solving for the baseline and then for the spherical harmonic coefficients.
We describe a procedure for this in \S\ref{sec:spotstar} below.

@@

In the following sections we discuss various approaches to solving the Doppler imaging problem for this dataset.
Unless otherwise noted, in all experiments we adopt a Gaussian prior for the spherical harmonic coefficients $\bvec{y}$ with mean $\boldsymbol{\mu}_\bvec{y} = \left(1 \quad\quad \bvec{0}\right)$ and covariance $\boldsymbol{\Lambda}_\bvec{y} = 10^{-4} \bvec{I}$, corresponding to a flat power spectrum prior.
When solving for the rest frame spectrum, we also place a Gaussian prior on $\bvec{s}$ with unit mean and covariance $\boldsymbol{\Lambda}_\bvec{s} = 10^{-3}\bvec{I}$.
These choices are arbitrary and primarily based on trial-and-error given knowledge of the true solutions.
We discuss strategies for choosing priors for real datasets in \S\ref{sec:discussion:priors} below.

@@

In our inference step, we again assume we have no knowledge of the map, the baseline, or either of the spectral components.
While our bilinear method (\S\ref{sec:spot_y1bs}) could in principle work here, we find that in practice it is very difficult to converge on the global optimum without either good priors or a good initial guess.
We therefore optimize our model using the nonlinear gradient-descent \texttt{Adam} optimizer \citep{Adam} within the \texttt{pymc3} probabilistic modeling framework \citep{Salvatier2016}.
Since we are no longer restricted to using Gaussian priors on the variables (a requirement of our bilinear method), we place a unit-mean Laplacian (sparsity) prior on the two spectra with scale parameter $b = 10^{-3}$ and bounded to the range $[0, 1]$.
%
We further place a Gaussian process (GP) prior on the spectra to enforce smoothness, adopting a squared exponential kernel with amplitude $10^{-1}$ and lengthscale $10^{-2}$~nm.
%
We also solve for a single scale parameter $r \sim \mathrm{Uniform}(0, 0.75)$, equal to the ratio of the continuum levels of the two spectra (whose true value is 0.5).
Finally, in order to enforce that the two component surface maps sum to unity (as described above), we optimize for the actual \emph{pixel} intensities of the photosphere map, on which we place a simple uniform prior on $[0, 1]$; the spot map is then simply the complement.
The spherical harmonic coefficients are then computed via a fast, linear spherical harmonic transform (SHT) operation at each step, as in \citet{Bartolic2021}.
In order to satisfy the Nyquist criterion, our pixel grid consists of twice as many pixels as spherical harmonic coefficients.%
\footnote{%
While it may seem strange to return to a discretized parametrization of the surface---after all the work we did to derive a closed-form solution in the spherical harmonic basis---our approach capitalizes on the strengths of both parametrizations: the ease of incorporating a nonnegativity prior in the pixel basis and the efficiency with which we can compute the model in the spherical harmonic basis.
Thanks to the linear relationship between these two bases, the transformation between them accrues virtually no overhead.}

@@

\section{Application to WISE 1049-5319B}
\label{sec:luhman16b}
%

Having shown that our model performs well on synthetic data, we now turn to a real dataset: the VLT/CRIRES observations of WISE 1049-5319B (Luhman 16B) presented in the Doppler imaging study of \citet{Crossfield2014}.
WISE 1049-5319B belongs to a binary brown dwarf system, which, at a distance of 2 pc, consists of the two closest known brown dwarfs to the Solar system.
Despite its proximity, WISE 1049-5319B is intrinsically dim and, with a projected rotational velocity $v \sin i = 26.1$ km/s, is at the slow end of what can be imaged using the traditional Doppler imaging technique.
Therefore, unlike those of classical Doppler imaging targets, the spectral lines of WISE 1049-5319B do not individually show large scale spot modulation (Figure~\ref{fig:spot}) from one epoch to the next; nor are the spectral templates for brown dwarfs particularly well understood.
Nonetheless, we chose to model WISE 1049-5319B specifically to show the power behind our new methodology, in particular its ability to learn both the surface map and the rest frame spectrum simultaneously from the entire dataset---not just from unblended spectral lines that have reliable templates.

In the original Doppler imaging study of this target, \citet{Crossfield2014} took 56 spectra of each object (the angular resolution was high enough to obtain individual spectra of both brown dwarfs) over the four CRIRES detectors in the range $2.288-2.345 \mu$m over a time span of 5 hours, with individual exposures of 300 s.
The 5 hour baseline was enough to capture one full rotation of WISE 1049-5319B, whose period was constrained to be about 4.9 hours from its photometric variability \citep{Gillon2013}.
These spectra were then downbinned to 14 epochs to boost the SNR, corrected for tellurics and spectrograph systematics, and continuum-normalized.
Each of the 14 spectra consist of 1,024 intensity measurements in each of 4 detectors, for a total of $57,344$ data points.
For more details on the data reduction, refer to \citet{Crossfield2014}.

@@

In order to produce Doppler images of WISE 1049-5319B, \citet{Crossfield2014} used brown dwarf template spectra from the BT-Settl library \citep{Allard2013}, finding that the spectrum with $\log_{10} g = 5.0$ and $T_\mathrm{eff} = 1450$ K provided the best match to the dataset.
This spectrum was then Doppler-shifted by the systemic radial velocity of the brown dwarf and broadened to match the spectrograph resolution.
This final template spectrum is shown as the blue curves in Figure~\ref{fig:luhman16b:template_spectra}, where we again masked regions where data was discarded.

@@

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/luhman16b_data_model.pdf} % 
        \caption{%
            Calibrated and tellurics-corrected CRIRES spectra of WISE 1049-5319B from \citet{Crossfield2014} taken at 14 epochs spanning a full rotation of the brown dwarf (black points).
            Our maximum \emph{a posteriori} solution is shown as the orange curves.
        }
        \label{fig:luhman16b:data_model}
    \end{centering}
\end{figure}
%
\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/luhman16b_spectra.pdf} % 
        \caption{%
            Unbroadened template spectra used to model the data in \citet{Crossfield2014} (blue curves). 
            These were used as a prior on the spectra inferred in this work (orange curves). 
            We infer a slightly different depth for some lines, as well as subtle changes in the continuum.
        }
        \label{fig:luhman16b:template_spectra}
    \end{centering}
\end{figure}

@@

As in \S\ref{sec:twospec}, our choice of parametrization and priors precludes the use of the fast bilinear approach outlined in \S\ref{sec:inverse}.
We therefore follow the same approach as in \S\ref{sec:twospec}, optimizing our log-likelihood function using \texttt{Adam} with a learning rate of $10^{-3}$.
Since we have a very good guess for the template spectrum, this optimization is much faster than in \S\ref{sec:twospec}, converging within $1,000$ steps.
Because of our efficient gradient-based implementation (\S\ref{sec:starry}), the entire optimization takes only about 3 minutes of runtime on a 2018 Macbook laptop.

\begin{figure}[t!]
    \begin{centering}
        \begin{minipage}[t]{0.49\linewidth}
            \centering
            \quad\quad\quad
            \textbf{Luger et al. (2021)}
        \end{minipage}
        \vspace{1em}
        \begin{minipage}[t]{0.49\linewidth}
            \centering
            \textbf{Crossfield et al. (2014)}
        \end{minipage}
        \includegraphics[width=0.495\linewidth]{figures/luhman16b_map.pdf} % 
        \includegraphics[width=0.462\linewidth]{static/luhman16b_crossfield.png} % 
        \caption{%
            Maximum \emph{a posteriori} surface maps of WISE 1049-5319B inferred from the spectra in Figure~\ref{fig:luhman16b:data_model}. 
            \emph{Left:} surface map inferred using the algorithm developed in this work. 
            \emph{Right}: surface map inferred in \citet{Crossfield2014}, reproduced from Figure~2 in that paper.
            We infer the same two dominant features as in \citet{Crossfield2014b}:
            a bright circumpolar spot, seen along the sub-observer meridian at $t = 0.0$ hr,
            and a dark, latitudinally elongated band prominent at $t = 2.4$ hr.
            While the exact latitude of the bright spot and the behaviour of the dark spot equatorward of $-30^\circ$ latitude differ betwee the two solutions, both are point estimates obtained under different choices of prior and from significantly different methods.
            We take their qualitative agreement to be a successful validation of our method.
        }
        \label{fig:luhman16b:maps}
    \end{centering}
\end{figure}

@@

The right panel of Figure~\ref{fig:luhman16b:maps} shows our principal result: the inferred solution for the surface brightness map of WISE 1049-5319B; compare to the right panel, showing the solution inferred by \citet{Crossfield2014}.
Broadly speaking, the two surface maps agree rather well: we infer the same circumpolar bright spot, of about the same size and at roughly the same phase, although the exact latitude is different.
We also infer a similar dark feature on the ``back'' side of the object (visible near the sub-observer point at $t = 2.4$ hours).
Bright equatorial regions also broadly match between both solutions.
There are some important differences, such as the behavior of the $t = 2.4$ hour dark region beyond $60^\circ$ of the visible pole.
In general, we find that different choices for our priors---in particular the regularization on the pixels and the template spectrum, the prior on the limb darkening coefficient, and the value of the inclination of the star---result in maps that differ somewhat.
We do, however, agree with \citet{Crossfield2014} in that the polar bright spot is the most robust of the inferred features, followed by the bottom portion of the dark spot seen at $t = 2.4$ hours.
Once again, since this is simply an \emph{optimization}, we are unable to place reliable uncertainties on our solutions.
These require a sampling scheme like MCMC (\S\ref{sec:discussion:priors}).

@@

\subsection{WISE 1049-5319B}
\label{sec:discussion:luhman}
